Parallelizing Linear Recurrent Neural Nets Over Sequence Length

### 2017-10-27 paper/main.tex
Medical section good to be included, we just need the final test score

% Experiments were performed on the N-MNIST \cite{orchard2015converting} dataset. N-MNIST captures the MNIST digit dataset \cite{lecun1998mnist} by panning
% an event driven camera over each digit. Each example in N-MNIST is a sequence of single pixel events (x, y, polarity, timestamp) where
% x and y each indicate a pixel position in [0, 33], polarity indicates whether the pixel was switching on or off, and timestamp is the time of the event in microseconds. Videos produced from the event data show that positive polarity events are often located on the leading edge of the digit motion and negative polarity events on the trailing edge. Between the 60,000 digits in the training set, N-MNIST contains approximately 250 million pixel events with sequence lengths ranging from 500 to 8000 and averaging 4000 events.

% We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with 256 units per layer and a two layer GILR-LSTM with 234 units per layer. The GILR-LSTM layer size was selected so that it had slightly fewer parameters than the LSTM. Each model transformed the incoming event position into a 40-dimensional embedding vector which was then combined with the polarity to produce a 41 dimensional input. Both models output a $2 \mathbf{x} 34^2$ matrix containing two future event location probability distributions conditioned on the polarity of the future event. The training algorithm only considers the probability distribution of the true future polarity and uses the cross entropy loss function. The Adam \cite{kingma2014adam} optimization algorithm and Glorot \cite{glorot2010understanding} initialization scheme were used. Training on a minibatch of size $b$ with sequence length $T$ consisted of uniformly sampling $b$ N-MNIST sequences and then extracting a single random $T$ element subsequence (and its 50 element ahead forecast) from the sequence. Sequences less than $T$ elements were padded out to length $T$, and there was a "burn-in" of 30 events at the start of each sequence where no predictions were made. An epoch was defined as a pass over as many pixel events as there are in the full training set.

% All experiments were performed using TensorFlow 1.0 \cite{abadi2016tensorflow} on a single Nvidia K80 GPU running for up to 18 hours. The LSTM model was computed using TensorFlow's dynamic\_rnn and BasicLSTMCell routines which are slower than but algorithmically similar to the cuDNN LSTM implementation.

% \subsection{Training performance}
% The speed of neural net training is not solely determined by the training throughput but also by the frequency of optimization steps. As an example, a batch method may process inputs at the same rate as a minibatch method, but the minibatch method will generally converge much faster on large datasets. On an infinite dataset, the batch method never takes a single optimization step regardless of its throughput. With this example in mind, it is clear that achieving fast training is a balancing act between training data throughput and optimization step latency and that training performance should be evaluated with training curves and not just throughput numbers such as time per epoch.

% The LSTM and GILR-LSTM offer differ latency and throughput tradeoffs. LSTM throughput depends only on minibatch size but LSTM latency depends on both minibatch size and sequence length. Experiments were conducted training the LSTM models with minibatch size 256 and sequence lengths 128, 256, 512, and 1024. Figure \ref{fig:learning_curves} shows the smaller sequence lengths led to faster initial learning but inferior final performance.

% The throughput and latency of the GILR-LSTM are both influenced by batch size and sequence length. Several GILR-LSTMs were trained with batch size 16 and sequence length 1024. This combination was chosen because of the nearly maximum throughput, the low latency, and the ease of building a minibatch given the distribution of sequence lengths in the data set. The 234 unit GILR-LSTM reaches a better training loss in roughly 4.5 hours than any of the LSTMs could reach in 18 hours. This experimental evidence indicates the GILR-LSTM is as powerful of a model as the LSTM and can be trained in a fraction of the time.

@@

\subsection{The Medical Setting}
A common application of RNNs is in medicine. In order to illustrate the
increased performance that the plr algorithm allows, we implement an GILR-LSTM
on the problem presented in \Chris{CITE}, the 2016 PhysioNet/Computing in
Cardiology Challenge 2016.  The aim of the Classification of Normal/Abnormal
Heart Sound Recordings challenge is to determine which of a set of EEGs are
produced by people with abnormal hearts.  The data is recorded from a variety of
sources and split into a training and testing set. The testing set is not
publicly available, but can be evaluated through a web portal. The top results
cover a large range of techniques, but generally rely on featurization of the
raw EEG records. We want to see if we are able to leverage our advantage at high
sequence-lenghts to train directly on the raw EEG. The EEGs are sampled at 2,000
Hz and range in length from several seconds to more than one hundred seconds,
giving a max sequence length of over 200,000. Since the testing server is not
equipped with a GPU, we cannot run PLR in the testing step. We resample the
records to 250 Hz so that we are able to evaluate the test results in a
reasonable amount of time, padding all records that are shorter than the longest.

The training set of data is unbalanced, with 75\% of the data corresponding to
normal hearts. We rebalance the data by oversampling the abnormal data by a factor
of three, including each abnormal heart three times.

Key points are:
\begin{itemize}
\item{With the fast training of the GILR-LSTM / PLR, we can train quickly on very
    long sequence-lengths}
\item{Allows us to get state-of-the-art results, even when performing very quickly
    and with little hyperparameter tuning}
\end{itemize}

NEED TO GET THOSE RESULTS!\Chris{Quite likely that the results won't be available
  for tomorrow, or that they won't be particularly impressive. We may have to
  scratch this entire section unfortunately}

@@

\end{thebibliography}

>>>>

% % Experiments were performed on the N-MNIST \cite{orchard2015converting}
% dataset. N-MNIST captures the MNIST digit dataset \cite{lecun1998mnist} by
% panning % an event driven camera over each digit. Each example in N-MNIST is a
% sequence of single pixel events (x, y, polarity, timestamp) where % x and y each
% indicate a pixel position in [0, 33], polarity indicates whether the pixel was
% switching on or off, and timestamp is the time of the event in
% microseconds. Videos produced from the event data show that positive polarity
% events are often located on the leading edge of the digit motion and negative
% polarity events on the trailing edge. Between the 60,000 digits in the training
% set, N-MNIST contains approximately 250 million pixel events with sequence
% lengths ranging from 500 to 8000 and averaging 4000 events.

% % We attempt to forecast 50 events ahead in N-MNIST using a two layer LSTM with
% 256 units per layer and a two layer GILR-LSTM with 234 units per layer. The
% GILR-LSTM layer size was selected so that it had slightly fewer parameters than
% the LSTM. Each model transformed the incoming event position into a
% 40-dimensional embedding vector which was then combined with the polarity to
% produce a 41 dimensional input. Both models output a $2 \mathbf{x} 34^2$ matrix
% containing two future event location probability distributions conditioned on
% the polarity of the future event. The training algorithm only considers the
% probability distribution of the true future polarity and uses the cross entropy
% loss function. The Adam \cite{kingma2014adam} optimization algorithm and Glorot
% \cite{glorot2010understanding} initialization scheme were used. Training on a
% minibatch of size $b$ with sequence length $T$ consisted of uniformly sampling
% $b$ N-MNIST sequences and then extracting a single random $T$ element
% subsequence (and its 50 element ahead forecast) from the sequence. Sequences
% less than $T$ elements were padded out to length $T$, and there was a "burn-in"
% of 30 events at the start of each sequence where no predictions were made. An
% epoch was defined as a pass over as many pixel events as there are in the full
% training set.

% % All experiments were performed using TensorFlow 1.0 \cite{abadi2016tensorflow}
% on a single Nvidia K80 GPU running for up to 18 hours. The LSTM model was
% computed using TensorFlow's dynamic\_rnn and BasicLSTMCell routines which are
% slower than but algorithmically similar to the cuDNN LSTM implementation.

% % \subsection{Training performance} % The speed of neural net training is not
% solely determined by the training throughput but also by the frequency of
% optimization steps. As an example, a batch method may process inputs at the same
% rate as a minibatch method, but the minibatch method will generally converge
% much faster on large datasets. On an infinite dataset, the batch method never
% takes a single optimization step regardless of its throughput. With this example
% in mind, it is clear that achieving fast training is a balancing act between
% training data throughput and optimization step latency and that training
% performance should be evaluated with training curves and not just throughput
% numbers such as time per epoch.

% % The LSTM and GILR-LSTM offer differ latency and throughput tradeoffs. LSTM
% throughput depends only on minibatch size but LSTM latency depends on both
% minibatch size and sequence length. Experiments were conducted training the LSTM
% models with minibatch size 256 and sequence lengths 128, 256, 512, and
% 1024. Figure \ref{fig:learning_curves} shows the smaller sequence lengths led to
% faster initial learning but inferior final performance.

% % The throughput and latency of the GILR-LSTM are both influenced by batch size
% and sequence length. Several GILR-LSTMs were trained with batch size 16 and
% sequence length 1024. This combination was chosen because of the nearly maximum
% throughput, the low latency, and the ease of building a minibatch given the
% distribution of sequence lengths in the data set. The 234 unit GILR-LSTM reaches
% a better training loss in roughly 4.5 hours than any of the LSTMs could reach in
% 18 hours. This experimental evidence indicates the GILR-LSTM is as powerful of a
% model as the LSTM and can be trained in a fraction of the time.

@@

\subsection{The Medical Setting}
A common application of RNNs is in medicine. In order to illustrate the
increased performance that the plr algorithm allows, we implement an GILR-LSTM
on the problem presented in , the 2016 PhysioNet/Computing in
Cardiology Challenge 2016.  The aim of the Classification of Normal/Abnormal
Heart Sound Recordings challenge is to determine which of a set of EEGs are
produced by people with abnormal hearts.  The data is recorded from a variety of
sources and split into a training and testing set. The testing set is not
publicly available, but can be evaluated through a web portal. The top results
cover a large range of techniques, but generally rely on featurization of the
raw EEG records. We want to see if we are able to leverage our advantage at high
sequence-lenghts to train directly on the raw EEG. The EEGs are sampled at 2,000
Hz and range in length from several seconds to more than one hundred seconds,
giving a max sequence length of over 200,000. Since the testing server is not
equipped with a GPU, we cannot run PLR in the inference. We resample the
records to 500 Hz so that we are able to evaluate the test results in a
reasonable amount of time, padding all records that are shorter than the longest.

The training set of data is unbalanced, with 75\% of the data corresponding to
normal hearts. We rebalance the data by oversampling the abnormal data by a factor
of three, including each abnormal heart three times. We train a network consisting
of three stacked GIRL-LSTM layers, each with 16 hidden units. Although reasonably
small, since the input is only one-dimensional 16 hidden units allows us to
develop representations. PLR allows us to train the network very rapidly over a
long sequence length, so we don't have to downsample and lose information or
spend lots of time developing methods to form features.
We were able to fit exactly to the training set after
a few hours of training, and on the unseen test data we obtained a score of
PLACEHOLDER.

\Chris{If we get good results (doesn't have to be best ever, but can be on par),
  then keep}
\Chris{If we don't get good results, or don't have results, we will need to kick
  this section}

@@

\bibitem[Goldberger et~al.(2000) Amaral, Glass, Hausdorff, Ivanov, Mark, Mietus,
    Moody, Peng, Stanley]{physiobank} Goldberger et~al. 
    \newblock PhysioBank, PhysioToolkit, and PhysioNet:
Components of a New Research Resource for Complex Physiologic
Signals.
\newblock In \emph{Circulation} 101(23):e215-e220; 2000 (June 13). PMID:
10851218; doi: 10.1161/01.CIR.101.23.e215
\end{thebibliography}
